---
title: "IMA Day 4 Notes"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---




## Getting Started

**Load packages required for exercises:**

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
```


## Motivating Example
```{r}
mydata <- data.frame(x = c(1:5))
mydata

sample_n(mydata, size=3)
sample_n(mydata, size=3, replace=TRUE)
```


### Exercise 1

```{r}
#Step 1: set up a 2 card deck
deck <- data.frame(card=c("fire", "no fire"))
deck

#Step 2: Randomly pick a card from the deck 17 times
suppressPackageStartupMessages(library(dplyr))
outcome <- sample_n(deck, size = 17, replace=TRUE)

#Step 3: Check out the results
table(outcome)

```

<br>
<br>

### Exercise 2

part a:

```{r}
#set the seed
set.seed(2000)
#initialize the for loop
Trials <- data.frame(fire = rep(0,500), nofire = rep(0,500))
#run the for loop
for(i in 1:500){
    #randomly pick a card from the deck 17 times
    outcome <- sample_n(deck, size = 17, replace=TRUE)

    #store the results in Trials
    Trials[i,] <- table(outcome)
}

#check out the first 6 experiment results
head(Trials)
```

part b:

```{r}
ggplot(Trials, aes(x=nofire)) + 
    geom_histogram(color="white")
```

part c:
```{r}
min(Trials$nofire)
max(Trials$nofire)
```
```{r}
table(Trials$nofire)
```


<br>
<br>


### Exercise 3

a. No. 14/17 is not a typical result
b. 3/500 = 0.006
c. Null hypothesis is false. We conclude that P.S has blindsight.


        
<br>
<br>

### Exercise 4
What if, instead, P.S. had chosen the non-burning house 10 out of 17 times.

a. Kind of Yes. Selects it
b. (56+50+ 34+10)/500 = 0.0316
c. Yes.

<br>
<br>
<br>
<br>



## Hypothesis Testing for Model Coefficients

```{r}
ps <- read.csv("https://www.macalester.edu/~ajohns24/data/PetStress.csv")
head(ps, 3)
```

### Exercise 5
b. On average, heart rate increases with age.
c. Symmetric Bell shaped distribution with mean zero.





<br>
<br>

### Exercise 6


part a:
```{r}
#plot
ggplot(ps, aes(x=Age,y=MaxHeartRate)) +
      geom_point() + 
      geom_smooth(method="lm")


#model
HeartMod1 <- lm(MaxHeartRate ~ Age, data=ps)
summary(HeartMod1)

```    
        
```{r}
confint(HeartMod1, level=0.95)
```

b. We don't have enough evidence to "reject" $H_{0}$ as 0 lies in the CI.

c. 0.3432
 t-value:how many standard errors away from 0 is $\hat{\beta_{1}}$
```{r}
0.2765/0.3432
```
It is relatively close to 0 by 68-95-99.7 rule
<br>
<br>

### Exercise 7
a. >0.05
```{r}
#Probability that it is more than 0.8* sd away from 0. (two sided)
1- 0.8*0.68

#one-sided
(1- 0.8*0.68)/2
```
c. It's not unlikely that we'd have observed such a strong increase in heart rate with age if in fact there were no relationship between these variables in the broader population.

d. Don't reject  $H_{0}$ . We don't have enough evidence to conclude that, on average, heart rate increases with age 

<br>
<br>


### Exercise 8
a
```{r}
#plot
ggplot(ps, aes(y=MaxHeartRate,x=MeanHeartRate)) +
      geom_point() + 
      geom_smooth(method="lm")


#model
HeartMod2 <- lm(MaxHeartRate ~ MeanHeartRate, data=ps)
summary(HeartMod2)
```
b.
```{r}
confint(HeartMod2, level=0.95)
```

$\hat{\beta_{1}}$ : Estimate = 1.6613, SE= 0.1062

c. The esitmate \hat{\beta_{1}} is t-value (15.642) times SE away from 0. This means if the null hypothesis were true, this would be quite unlikely (w.r.t the significance level $\alpha = 0.05$).

d. p-value 

```{r}
#one-sided
(1- 15.642*0.68)/2
```

The probability that this data is not as extreme if our null hypothesis is true is p-value < 2.2e-16

e. Thus, we reject the null hypothesis that there is no co-relation between person's maximum heart rate to their mean heart rate.

<br>
<br>


### Exercise 9

```{r}
#plot
ggplot(ps, aes(y=MaxHeartRate,x=MeanHeartRate, color = Condition)) +
      geom_point() + 
      geom_smooth(method="lm")
#model
HeartMod3 <- lm(MaxHeartRate ~ MeanHeartRate + Condition , data=ps)
summary(HeartMod3)
```

```{r}
confint(HeartMod3, level=0.95)
```
b. $\hat{\beta_{0}}$: MaxHeartBeat of a person when their MeanHeartRate = 0 and subject is in controlled environment.
$\hat{\beta_{1}}$: Change in MaxHeartBeat when MeanHeartBeat change by $\hat{\beta_{1}}$ in controlled environment.

$\hat{\beta_{2}}$ The amount by which MaxHeartBeat differs from when they are in controlled environment of a person when their MeanHeartRate = 0 and they are with a friend

$\hat{\beta_{3}}$ The amount by which MaxHeartBeat differs from when they are in controlled environment of a person when their MeanHeartRate = 0 and they are with a pet

c. While Controlling for the Condition parameter, is MeanHeartRate and MAxHeartRate associated with each other significantly

The esitmate \hat{\beta_{1}} is t-value (15.642) times SE away from 0 while contolling for the Condition parameter. 

p-value < 2e-16

Controlling for the Condition Paramter, we have significant evidence to reject the null hypothesis i.e we reject the hypothesis that there is no co-relation between MeanHeartRate and MaxHeartRate.

d.ConditionF
Null hypothesis would state there is difference in MaxHeartRate for same MeanHeartRate as you change environment from controlled to Friend present.

The estimate \hat{\beta_{2}} is t-value (-0.160) times SE away from 0 when meanheartrate is same as you change environment from controlled to Friend present.

p-value = 0.873/2  - Probability that  assuming $H_{0}$. 

Not enough evidence to reject the null hypothesis

e. Reject Null Hypothesis


<br>
<br>

### Exercise 10
```{r}
interactionMod <- lm(MaxHeartRate ~ MeanHeartRate*Condition, ps)
summary(interactionMod)
```

```{r}
confint(interactionMod, level=0.95)
```
MeanHeartRate:ConditionF - Not significant - Don't reject Null Hypothesis
MeanHeartRate:Conditionp - Significant - Reject Null Hypothesis

<br>
<br>
<br>
<br>


## Potential Errors in Hypothesis Testing

### Exercise 11

```{r}
dd <- c(0.01, 0.03, 0.59, 0.02, 0.55, 0.09, 1.36, 0.61, 0.18, 0.06)
mean(dd)
```
```{r}
curve(dnorm(x, mean=0.2, sd=0.05))
```

```{r}
pnorm(0.35, mean=0.2, sd=0.05, lower=FALSE)
```
c. Reject  $H_{0}$ . We have statistically significant evidence that the driver is not sober.
<br>
<br>

### Exercise 12
a. Type 1 error: Conclude that a sober driver is drunk.
b. Type II Error: Drunk driver is sober


<br>
<br>

### Exercise 13


```{r}
qnorm(0.95, mean=0.2, sd=0.05)
```

```{r}
sober <- rnorm(1000, mean=0.2, sd=0.05)
sum(sober> 0.2822)/10
```

<br>
<br>

### Exercise 14

```{r}
BAC8 <- rnorm(1000, mean=(0.2+ 4*0.08), sd=(0.05+0.08)**2)
```

```{r}
sum(BAC8 < 0.2822)/10
```

```{r}
BAC15 <- rnorm(1000, mean=(0.2+ 4*0.15), sd=(0.05+0.15)**2)
```

```{r}
sum(BAC15 <= 0.2822)/10
```
<br>
<br>

### Exercise 15



