---
title: "IMA Day 3 Notes"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---




## Getting Started

**Load packages required for exercises:**

```{r}
suppressPackageStartupMessages(library(rvest))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tidyr))
```


## Scraping & Wrangling Data {-}

**Scrape data from [Wiki](https://en.wikipedia.org/wiki/Current_members_of_the_United_States_House_of_Representatives)**

```{r}
#load the rvest library (if it's not already, install it first!)
suppressPackageStartupMessages(library(rvest))

#copy the Wiki url
url <- "https://en.wikipedia.org/wiki/Current_members_of_the_United_States_House_of_Representatives"

#scrape ALL tables from the url
allTables <- read_html(url) %>% 
    html_nodes("table") %>% 
    html_table(fill=TRUE)

#the 7th table corresponds to "Voting members by state".  store this as "houseDirty"
houseDirty <- allTables[[7]]

#check out the head of houseDirty
names(houseDirty)

head(houseDirty, 3)

date()

write.csv(houseDirty, "D:/Math2IndustryBootCamp2017/house.csv", row.names=FALSE)


```


<br>
<br>

### Exercise 1

```{r}
#part a
mytext <- "This is my text"
gsub("s", "z", mytext)
gsub(" ", "", mytext)
substr(mytext, 1, 3)
substr(mytext, 4, 7)

#part b
mynumber <- "2017"
class(mynumber)
mynewnumber <- as.numeric(mynumber)
mynewnumber + 2
class(mynewnumber)

```

<br>
<br>

### Exercise 2

```{r}
#load the dplyr library for some wrangling functions
suppressPackageStartupMessages(library(dplyr))

#take out the third column (the blank "Party" variable)
house <- houseDirty[,-3]

#remove the spaces from the variable names
names(house) <- gsub(" ", "", names(house))

#check the class of the Assumedoffice variable 
class(house$Assumedoffice)

#in 1 set of code: 
#remove the "*" from the Assumedoffice measurements
#convert Assumedoffice to numeric
#define an "Age" variable
house$Assumedoffice <- substr(house$Assumedoffice,1,4) %>% as.numeric(house$Assumedoffice) 

house <- house %>% mutate(Age = 2017 - house$Bornin)
#confirm that your house is consistent with the results in the notes


```

<br>
<br>

## Sampling Variability
```{r}
suppressPackageStartupMessages(library(ggplot2))
mean(house$Age, na.rm=TRUE)
sd(house$Age, na.rm=TRUE)
ggplot(house, aes(x=Age)) + 
    geom_histogram(color="white") + 
    labs(title="POPULATION distribution")
```


### Exercise 3

```{r}
#choose your own random number seed
set.seed(37)
#take a sample of 16 members
sample1 <- sample_n(house,size = 16)

#check the properties
head(sample1)
dim(sample1)
mean(sample1$Age)#-mean(house$Age)
sd(sample1$Age)#- sd(house$Age)
```


part b:
```{r}
ggplot(sample1, aes(x=Age)) + 
    geom_histogram(color="white", binwidth=5) + 
    labs(title="SAMPLE Distribution (n=16)") + 
    lims(x=c(30,90), y=c(0,80))

```    

part c:
```{r}

```




<br>
<br>

### Exercise 4

```{r}
#take another sample

```




<br>
<br>


## Sampling Distributions



<br>

Sample Size   Shape         Min $\overline{y}$  Mean $\overline{y}$   Max $\overline{y}$  St Dev of $\overline{y}$ 
------------- ------------- ------------------- --------------------- ------------------- --------------------------
$n = 16$      Bell          50                  58.572                66.75               2.5
$n = 64$      Smaller Bell  55.10               58.49                 63.48               1.24
$n = 100$     Smallest Bell 55.77               58.55                 61.64               0.934


      
<br>
<br>

### Exercise 5

```{r}
#set the seed
set.seed(2)

#initialize the for loop
simulation <- data.frame(n16=rep(0,1000), n64=rep(0,1000), n100=rep(0,1000))

#run the for loop
for (i in 1:1000){
  
#take a sample of 16 members
  sample16 <- sample_n(house,size = 16)
  simulation$n16[i] = mean(sample16$Age)
  sample64 <- sample_n(house,size =64)
  simulation$n64[i] = mean(sample64$Age)
  sample100 <- sample_n(house,size =100)
  simulation$n100[i] = mean(sample100$Age)
}


```   

```{r}
min(simulation$n100)
mean(simulation$n100)
max(simulation$n100)
sd(simulation$n100)

```
```{r}
ggplot(simulation, aes(x=n100)) + 
    geom_histogram(color="white", binwidth=2) + 
    labs(title="SAMPLING Distribution") + 
    lims(x=c(45,75)) + 
    geom_vline(xintercept=58.54734, color="red")
```


<br>
<br>

### Exercise 6

```{r}

```


    


<br>
<br>

### Exercise 7

```{r}

```


<br>
<br>

### Exercise 8

```{r}

```


<br>
<br>

### Exercise 9

```{r}
#follow directions
head(simulation, 3)
##     n16   n64  n100
## 1 57.81 59.55 61.29
## 2 57.06 58.72 59.45
## 3 57.19 59.95 59.34
dim(simulation)

suppressPackageStartupMessages(library(tidyr))
simulation <- simulation %>% 
    gather(key=SampleSize, value=SampleMean)
head(simulation, 3)
dim(simulation)
```


<br>
<br>

### Exercise 10

```{r}
#density plot by group

```


<br>
<br>

### Exercise 11

```{r}
sd(house$Age, na.rm=TRUE)/sqrt(16)
sd(house$Age, na.rm=TRUE)/sqrt(64)
sd(house$Age, na.rm=TRUE)/sqrt(100)
## [1] 10.84
## [1] 10.84
## [1] 10.84
```




<br>
<br>


## Regression Models

### Exercise 12
```{r}
PopMod <- lm(Age ~ Assumedoffice, data=house)

#obtain just the coefficients from this model
PopMod$coef
##   (Intercept) Assumedoffice 
##     1468.9169       -0.7024

ggplot(house, aes(x=Assumedoffice, y=Age)) + 
    geom_point() + 
    geom_smooth(method="lm")
```

part a:
```{r}
#set the seed
set.seed(20)
simulation2 <- data.frame(beta0=rep(0,1000), beta1=rep(0,1000))

#initialize the for loop
for (i in 1:1000){
  newsamp16 <- sample_n(house, size=100)
  #dim(newsamp16)
  SampMod1 <- lm(Age ~ Assumedoffice, data=newsamp16)
  simulation2$beta0[i] = SampMod1$coef[1]
  simulation2$beta1[i] = SampMod1$coef[2]
}

#run the for loop

```
    
part b:

```{r}
#construct a histogram of the 1000 beta0 estimates
ggplot(simulation2, aes(x=beta0)) + 
    geom_histogram() 
    
#construct a histogram of the 1000 beta1 estimates
ggplot(simulation2, aes(x=beta1)) + 
    geom_histogram() 

```

part c:
```{r}
#approximate the standard errors
mod <- lm(mod <- lm(Age ~ Assumedoffice, data=house))
summary(mod)

summary(simulation2$beta0)
sd(simulation2$beta0)

summary(simulation2$beta1)
sd(simulation2$beta1)
```


<br>
<br>

### Exercise 13

part a:
```{r}
#set up the plotting frame and store it as g1
g1 <- ggplot(house, aes(x=Assumedoffice,y=Age)) + 
    lims(y=c(40,100), x=c(1970,2016))

for(i in 1:20){
    #take a sample of 16 members
    samp <- sample_n(house, size=16)

    #add the sample model line to g1
    g1 <- g1 + 
        geom_smooth(data=samp, aes(x=Assumedoffice,y=Age), 
            method="lm", color="gray", se=FALSE, fullrange=TRUE)
}

#plot the 20 sample lines with the population line (red)
g1  + geom_smooth(data=house, aes(x=Assumedoffice,y=Age), 
    method="lm", color="red", se=FALSE) 
```

part b:
```{r}
#set up the plotting frame and store it as g1
g1 <- ggplot(house, aes(x=Assumedoffice,y=Age)) + 
    lims(y=c(40,100), x=c(1970,2016))

for(i in 1:20){
    #take a sample of 16 members
    samp <- sample_n(house, size=100)

    #add the sample model line to g1
    g1 <- g1 + 
        geom_smooth(data=samp, aes(x=Assumedoffice,y=Age), 
            method="lm", color="gray", se=FALSE, fullrange=TRUE)
}

#plot the 20 sample lines with the population line (red)
g1  + geom_smooth(data=house, aes(x=Assumedoffice,y=Age), 
    method="lm", color="red", se=FALSE) 
```


    
    

    
    

<br>
<br>


## The Central Limit Theorem

### Exercise 14



<br>
<br>


## Extra: Writing Functions






<br>
<br>
<br>


