---
title: "IMA Day 1 Notes"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---




## Getting Started

**Load packages required for exercises:**

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(choroplethr))
suppressPackageStartupMessages(library(choroplethrMaps))
```


**Load the data**

```{r}
#this uses the readr library:
elect <- read.csv("https://raw.githubusercontent.com/tonmcg/County_Level_Election_Results_12-16/master/US_County_Level_Presidential_Results_08-16.csv")
```



<br>
<br>


## Pre-Bootcamp Review

### Exercise 1: Explore the structure of the data

```{r}
#Check out the first rows of `elect`.  What are the cases?  The variables?    
head(elect,1)
# cases = counties: election data
# variables = columns
    
#How much data do we have?    
dim(elect)
    
#What are the names of the variables?    
names(elect)
# or colnames(elect)
    
#How can we subset just the `dem_2016` data?    
head(elect$dem_2016)
 

#How can we filter out only the counties in which more than 25,000 people voted in 2016?
library(dplyr)
elect %>% filter(total_2016 >= 25000) %>% head(2)

```
    


<br>
<br>

### Exercise 2: Create new variables


```{r}
#start with 1 variable to get the hang of it
elect %>% 
    mutate(perdem_2016 = (dem_2016/total_2016))
        
#now 2 variables    
#elect %>% mutate(perrep_2016 = (rep_2016/total_2016))
#elect %>% mutate(winrep_2016 = )

#now 3.  (why do we put winrep_2016 in its own row?!)


#finally, add all 2012 and 2016 variables and store as 'elect'
elect <- elect %>% mutate(perdem_2016 = (dem_2016/total_2016), perrep_2016 = (gop_2016/total_2016), winrep_2016 = (dem_2016 < gop_2016), perdem_2012 = (dem_2012/total_2012), perrep_2012= (gop_2012/total_2012), winrep_2012 = (dem_2012 < gop_2012))

head(elect)

```
        


<br>
<br>

### Exercise 3: Explore the win column

```{r}
#Construct a table (a numerical summary) of the number of counties that Trump won/lost
table(elect$winrep_2016)
#Construct a bar chart (a visual summary) of this variable.    
elect$winrep_2016
```





<br>
<br>

### Exercise 4: Visually explore vote percentages

```{r}
#histogram
ggplot(elect, aes(x=perrep_2016)) + geom_histogram(color="white",fill = 'blue')
# We see typical behaviour, max min

#density plot
ggplot(elect, aes(x=perrep_2016)) + geom_density(color="blue")
# has an area = 1, relative stuff in [0,1]

```



<br>
<br>

### Exercise 5: Numerically summarize vote percentages

```{r}
#calculate the mean Trump support
mean(elect$perrep_2016)

#calculate the median Trump support
median(elect$perrep_2016)

#calculate the variance in Trump support from county to county
var(elect$perrep_2016)

#calculate the standard deviation in Trump support from county to county
sd(elect$perrep_2016)
#use the 68-95-99.7 Rule to approximate the range of from the 2.5th to 97.5th percentiles of  Trump support
mean(elect$perrep_2016) - 2*sd(elect$perrep_2016)
mean(elect$perrep_2016) + 2*sd(elect$perrep_2016)
#calculate the exact 2.5th and 97.5th percentiles 
quantile(elect$perrep_2016, c(0.025,0.975))
```


<br>
<br>
<br>
<br>


## Explaining Variability + Wrangling


**Load the data:**

```{r warning=FALSE}
#load demographic data from the choroplethr library
data("df_county_demographics")
head(df_county_demographics,3)

#load RedBluePurple data
RedBlue <- read.csv("https://www.macalester.edu/~ajohns24/Data/RedBluePurple.csv")
head(RedBlue,3)
```



<br>
<br>

**Join / merge these data tables into a single data set:**

```{r}
#define "region" variable
elect <- elect %>% mutate(region=fips_code)

#join elect and df_county_demographics (from dplyr)
fulldata <- left_join(df_county_demographics, elect)
#join the new fulldata with RedBlue
fulldata <- left_join(fulldata, RedBlue)


#check out the names of fulldata
names(fulldata)

```


<br>
<br>
<br>
<br>


## Visualizing Relationships

### Exercise 6: Scatterplots of 2 quantitative variables


```{r}
#just a graphics frame
ggplot(fulldata, aes(y=perrep_2016, x=perrep_2012))
#adding a scatterplot layer
ggplot(fulldata, aes(y=perrep_2016, x=perrep_2012)) + geom_point()

#another predictor
ggplot(fulldata, aes(y=perrep_2016, x=median_rent)) + 
    geom_point()

#another predictor
ggplot(fulldata, aes(y=perrep_2016, x=percent_white)) + 
    geom_point()
```
    
    
   
<br>
<br>

### Exercise 7: Side-by-side plots of 1 quantitative variable vs 1 categorical variable


```{r}
#density plots by group
ggplot(fulldata, aes(x=perrep_2016, fill=StateColor)) + 
    geom_density(alpha=0.5)+
  scale_fill_manual(values = c('blue' ,'purple','red'))

#to see better: add transparency
ggplot(fulldata, aes(x=perrep_2016, fill=StateColor)) + 
    geom_density(alpha=0.5)


#to see better: split groups into separate plots


#violins instead


#boxes instead


```



<br>
<br>


### Exercise 8: Scatterplots of 1 quantitative variable vs 1 categorical & 1 quantitative variable


```{r}
#scatterplot: id groups using color


#scatterplot: id groups using shape


#scatterplot: split/facet by group


```



<br>
<br>


### Exercise 9: Side-by-side plots of 1 quantitative variable vs 2 categorical variables

```{r}
#density plots by group combo


#switch faceting order


```

<br>
<br>

### Exercise 10: Plots of 3 quantitative variables

```{r}
#scatterplot: represent third variable using color


#scatterplot: discretize the third variable into 2 groups & represent with color


#scatterplot: discretize the third variable into 2 groups & facet by group


```


    
<br>
<br>


### Exercise 11: Maps!

```{r}
#Construct maps using the choroplethrMaps library

#a map of Trump support


#a map of Trump wins


#a map of percent white


```




<br>
<br>

<br>
<br>



## Linear Regression Models

### Exercise 12: Models with 1 quantitative predictor


```{r}
#visualization
ggplot(fulldata, aes(x=percent_white, y=perrep_2016)) + 
    geom_point(alpha=0.25) + 
    geom_smooth(method="lm")

#model
model1 <- lm(perrep_2016 ~ percent_white, data=fulldata)
summary(model1)

# \beta_{0} gives trump_spport when there is no white population in the total population
# \beta_{1} gives amount increase in trump_support when white population changes by 1
```


<br> 
<br>

### Exercise 13: Models with 1 categorical predictor

```{r}
#plot the relationship
ggplot(fulldata, aes(x=perrep_2016, fill=StateColor)) + 
    geom_density(alpha=0.5)

```


```{r}
#detach/unload libraries (a bug)
detach(package:choroplethr)
detach(package:acs)
detach(package:plyr)
```

```{r}
#calculate the group means
fulldata %>%                                        #start with full data
    group_by(StateColor) %>%                        #group it by state color
    summarize(means=mean(perrep_2016, na.rm=TRUE))  #calculate the mean by state
```



```{r}
#construct the model
model2 <- lm(perrep_2016 ~ StateColor, data=fulldata)
summary(model2)

```



<br>        
<br>

### Exercise 14: Models with 1 quantitative predictor & 1 categorical predictor


```{r}
#visualize the relationship
ggplot(fulldata, aes(x=percent_white, y=perrep_2016, color=StateColor)) + 
    geom_point(alpha=0.25) + 
    geom_smooth(method="lm") + 
    facet_wrap(~ StateColor)


#construct the model

model3 <- lm(perrep_2016 ~ percent_white + StateColor, data=fulldata)
summary(model3)
```



<br>
<br>


### Exercise 15: Models with 2 quantitative predictors

```{r}
#visualize the relationship

#construct the model


```



<br> 
<br>

### Exercise 16: Models with 2 categorical predictors

```{r}
#create income brackets

```

```{r}
#visualize the relationship

#construct the model

```




<br>
<br>
<br>
<br>


