---
title: "Homework 5: Classification with Logistic Regression"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---


**Load packages:**

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(mosaic))
```

**Load data**

```{r}
diag <- read.csv("https://www.macalester.edu/~ajohns24/data/breastcancer2.csv")
head(diag)
```
```{r}
dim(diag)
```


### Exercise 1
a.
```{r}
mod1 <- glm(Class ~ Thick, diag, family = 'binomial')
summary(mod1)
```



```{r}
diag$probPredictions <- mod1$fitted.values

diag %>% ggplot(aes(y = probPredictions, x = Thick))+ stat_smooth(se=FALSE) 
```

c.
```{r}
100*(exp(0.94611) -1)
```
157.56 Percent increase in odds of getting a tumor for an increase in thichness by 1

Yes. p-value <0.05

0.875
Malignant
<br>
<br>

### Exercise 2
a. p= 0.5: odds are 1 for malignant tumor. log(odds) = 0
b. ~5.5 visually
```{r}
5.18/0.94
```

c. If thickness is greater than 5.5, conclude the tumor is malignant. If thickness is lesser than 5.5, conclude the tumor is benign.

<br>
<br>


### Exercise 3
    
```{r}
ggplot(diag, aes(x=Thick, fill=Class)) + 
    geom_density(alpha=0.5) + 
    geom_vline(xintercept=5.48)
```
a. sensitivity = P(classify a malignant case as malignant)
   specificity = P(classify a benign case as benign)
  
b. 

    
```{r}
#diagnose by the thickness cut-off
table(diag$Class, diag$Thick >= 5.48)

#diagnose by the probability cut-off
table(diag$Class, mod1$fitted >= 0.5)  
```
```{r}
sensitivity1 = 163/(163+75)
sensitivity1

specificity1 = 424/(424 + 19)
specificity1

```

```{r}
suppressPackageStartupMessages(library(boot))
cv.err <- cv.glm(diag, mod1, K= 681)
cv.err$delta
```



<br>
<br>

### Exercise 4

Cutoff = 0.2
```{r}

t = ((log(0.2/0.8))+5.18)/0.94
t
#diagnose by the thickness cut-off
table(diag$Class, diag$Thick >= t)

#diagnose by the probability cut-off
table(diag$Class, mod1$fitted >= 0.2)  
```

```{r}
sensitivity2 = 207/(207+31)
sensitivity2

specificity2 = 341/(341 + 102)
specificity2

```
```{r}
ggplot(diag, aes(x=Thick, fill=Class)) + 
    geom_density(alpha=0.5) + 
    geom_vline(xintercept=t)
```

Lowering cut off  --> Increase in sensitivity and decrease in specificity
<br>
<br>

### Exercise 5

```{r}
ggplot(diag, aes(x=jitter(Thick,0.75), y=jitter(USize,0.75), color=Class)) + 
    geom_point() + 
    labs(x="Thick", y="USize")
```


```{r}
mod2 <- glm(Class ~ Thick * USize, diag, family="binomial")
summary(mod2)
```
```{r}
#set up a prediction function
probFun <- makeFun(mod2)

#set up a grid of Thick & USize pairs at which to make predictions
griddata <- expand.grid(list(Thick=seq(0,10,length=50),USize=seq(0,10,length=50)))
griddata <- griddata %>% 
    mutate(prob=probFun(Thick=Thick, USize=USize)) %>%  #calculate the probability of malignancy at each pair
    mutate(Diagnosis = (prob > 0.5))                    #classify each pair

#plot the classification status (ie the diagnosis) at each pair
ggplot(griddata, aes(x=Thick, y=USize) ) +
    geom_raster(aes(x=Thick, y=USize, fill=Diagnosis))
```

```{r}
table(diag$Class, mod2$fitted >= 0.5)
```

```{r}
sensitivity3 = 220/(220+18)
sensitivity3

specificity3 = 429/(429+14)
specificity3
```

```{r}
mod3 <- glm(Class ~ Thick + USize, diag, family="binomial")
summary(mod3)
```

```{r}
#set up a prediction function
probFun3 <- makeFun(mod3)

#set up a grid of Thick & USize pairs at which to make predictions
griddata1 <- expand.grid(list(Thick=seq(0,10,length=50),USize=seq(0,10,length=50)))
griddata1 <- griddata1 %>% 
    mutate(prob=probFun3(Thick=Thick, USize=USize)) %>%  #calculate the probability of malignancy at each pair
    mutate(Diagnosis = (prob > 0.5))                    #classify each pair

#plot the classification status (ie the diagnosis) at each pair
ggplot(griddata1, aes(x=Thick, y=USize) ) +
    geom_raster(aes(x=Thick, y=USize, fill=Diagnosis))
```



