---
title: "Homework 3: Confidence Intervals"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---


**Load some handy packages:**

```{r, warning=FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
```


<br>
<br>

## Constructing Confidence Intervals


```{r}
sleep = read.csv("https://www.macalester.edu/~ajohns24/Data/SleepStudy2.csv")
head(sleep, 2)
```
<br>
<br>

### Exercise 1


```{r}
n = nrow(sleep)
mean(sleep$Difference)
sd(sleep$Difference)/sqrt(n)
```

1c 95%, 5%
<br>
<br>

### Exercise 2

2a. 95%  
2b
```{r}
mean(sleep$Difference) + 2*sd(sleep$Difference)/sqrt(n)
mean(sleep$Difference) - 2*sd(sleep$Difference)/sqrt(n)
```
2c.
```{r}
t.test(sleep$Difference, conf.level=0.95)
```

    
<br>
<br>

### Exercise 3
Yes. 0 falls within the confidence interval. Thus, our data don't signifcantly support the claim that that the average reaction time
after 2 days of sleep deprivation is greater (slower) than after 0 days of sleep deprivation.



<br>
<br>

## Interpreting Confidence Level
 

```{r}
#attach the library
suppressPackageStartupMessages(library(gapminder))

#load the data
data(gapminder)

#examine the codebook
?gapminder
```


<br>

### Exercise 4
```{r}
#4a. Number of cases
nrow(gapminder)
#4b For what years do we have data?
unique(gapminder$year)

#4c For how many different countries do we have data?
length(unique(gapminder$country))
```
4a. Each case represent is Country's life expectancy and gpa per capita for a particular year along with the continent it belongs to.




<br>

### Exercise 5
5a
```{r}
gapminder%>% ggplot(aes(y = lifeExp, x= gdpPercap)) + geom_point() + geom_smooth(method = 'lm') 
```


5b. Data points are not independent. Afghanistan's lifeExp (GDP) is dependent on previous year's lifeExp (GDP).  Also, residuals are not normally distributed with mean zero. (lack of fit)
```{r}
mod <- lm( lifeExp ~ gdpPercap, gapminder )
ModResults <- data.frame(observed = gapminder$lifeExp, predicted = mod$fitted.values, residual = mod$residuals)
ggplot(ModResults, aes(sample=residual)) + 
    geom_qq()
```


```{r}
gapminder%>% ggplot(aes(y = lifeExp, x= log(gdpPercap))) + geom_point() + geom_smooth(method = 'lm') 
```

```{r}
gap2007 <- gapminder %>% 
    filter(year==2007)

gap2007 %>% ggplot(aes(x = log(gdpPercap), y = lifeExp, color = continent, size = pop)) + geom_point()
```


<br>
<br>


### Exercise 6

```{r}
ggplot(gap2007, aes(x=lifeExp)) + 
    geom_histogram(color="white")

```

```{r}
mean(gap2007$lifeExp)

#set the seed
set.seed(39)

#take a sample
samp1 <- sample_n(gap2007, size=25)

#calculate the sample mean & sd
mean(samp1$lifeExp)
## [1] 72.5
sd(samp1$lifeExp)
## [1] 9.724
```
```{r}
mean(samp1$lifeExp)+ 2*sd(samp1$lifeExp)/sqrt(25)
mean(samp1$lifeExp)- 2*sd(samp1$lifeExp)/sqrt(25)

```

```{r}
t.test(samp1$lifeExp, conf.level=0.95)
```
6c. No. True mean falls outside our CI

<br>
<br>

### Exercise 7

```{r}
#set the seed
set.seed(90)

#initialize the for loop
CIsimulation <- data.frame(lower=rep(0,100), upper=rep(0,100))


#run the for loop
for(i in 1:100){
    samp <- sample_n(gap2007, size=25)
    CIsimulation[i,] <- t.test(samp$lifeExp, conf.level=0.95)$conf.int
}

```

7b. 
```{r}
CIsimulation <- CIsimulation %>% mutate(experiment = c(1:100), cover = (lower < 67.00742) & (upper > 67.00742))
```

```{r}
ggplot(CIsimulation, aes(y=experiment, x=lower, color=cover)) + 
    geom_segment(aes(x=lower, xend=upper, y=experiment, yend=experiment)) + 
    labs(x="confidence interval") + 
    lims(x=c(50,80))
```
```{r}
CIcovered <- CIsimulation %>% filter(cover == TRUE)
nrow(CIcovered)
```

<br>
<br>

### Exercise 8

```{r}
#set the seed
set.seed(90)

#initialize the for loop
CIsimulation <- data.frame(lower=rep(0,100), upper=rep(0,100))


#run the for loop
for(i in 1:100){
    samp <- sample_n(gap2007, size=75)
    CIsimulation[i,] <- t.test(samp$lifeExp, conf.level=0.95)$conf.int
}


CIsimulation <- CIsimulation %>% mutate(experiment = c(1:100), cover = (lower < 67.00742) & (upper > 67.00742))

ggplot(CIsimulation, aes(y=experiment, x=lower, color=cover)) + 
    geom_segment(aes(x=lower, xend=upper, y=experiment, yend=experiment)) + 
    labs(x="confidence interval") + 
    lims(x=c(50,80))

```

Larger sample size = narrower CI

<br>
<br>

### Exercise 9

Narrower but inaccurate as it is only 68% CI
```{r}
#set the seed
set.seed(90)

#initialize the for loop
CIsimulation <- data.frame(lower=rep(0,100), upper=rep(0,100))


#run the for loop
for(i in 1:100){
    samp <- sample_n(gap2007, size=25)
    CIsimulation[i,] <- t.test(samp$lifeExp, conf.level=0.68)$conf.int
}


CIsimulation <- CIsimulation %>% mutate(experiment = c(1:100), cover = (lower < 67.00742) & (upper > 67.00742))

ggplot(CIsimulation, aes(y=experiment, x=lower, color=cover)) + 
    geom_segment(aes(x=lower, xend=upper, y=experiment, yend=experiment)) + 
    labs(x="confidence interval") + 
    lims(x=c(50,80))

```


9c

```{r}
#set the seed
set.seed(90)

#initialize the for loop
CIsimulation <- data.frame(lower=rep(0,100), upper=rep(0,100))


#run the for loop
for(i in 1:100){
    samp <- sample_n(gap2007, size=25)
    CIsimulation[i,] <- t.test(samp$lifeExp, conf.level=0.9997)$conf.int
}


CIsimulation <- CIsimulation %>% mutate(experiment = c(1:100), cover = (lower < 67.00742) & (upper > 67.00742))

ggplot(CIsimulation, aes(y=experiment, x=lower, color=cover)) + 
    geom_segment(aes(x=lower, xend=upper, y=experiment, yend=experiment)) + 
    labs(x="confidence interval") + 
    lims(x=c(50,80))

```



Increasing % confidence means we are more likely to get an interval which captures the true mean but the size of the interval gets bigger. 
The greater the confidence level, the greater our confidence that the CI covers the population parameter but the wider (and less useful) the interval. Why 95%? Tradition.
<br>
<br>



## Confidence & Prediction Intervals for Regression Models

```{r}
PopMod <- lm(lifeExp ~ log(gdpPercap), data=gap2007)
PopMod$coef
summary(PopMod)
```


### Exercise 10
```{r}
set.seed(60)
samp50 <- sample_n(gap2007, size=50)
sampMod <- lm(lifeExp ~ log(gdpPercap), data=samp50)
summary(sampMod)
```
10a. $R^2 = 0.6687$ which means 66.87% of the variability in the data can be explained by our model

10b. $\hat{\beta_0} = 3.93$ and $\hat{\beta_1} = 7.3636$ 

10c. 95% CI
```{r}
#$\hat{\beta_0}$
3.93 + 2*6.5152
3.93 - 2*6.5152

7.3636 + 2*0.7482
7.3636 - 2*0.7482

```
```{r}
confint(sampMod, level=0.95)
```
Yes, the interval for the GDP coef is above 0
<br>
<br>

### Exercise 11

```{r}
3.93 + 7.3636 *log(42951.65)
```
```{r}
gap2007 %>% filter(country=="United States")
```
```{r}
predict(sampMod, newdata=data.frame(gdpPercap=42951.65),

interval="confidence", level=0.95)
```


```{r}
predict(sampMod, newdata=data.frame(gdpPercap=42951.65),
    interval="prediction", level=0.95)
```


<br>
<br>

### Exercise 12

```{r}
samp50 %>% ggplot(aes(x = log(gdpPercap), y = lifeExp)) + geom_point() + geom_smooth(method = 'lm')
```


```{r}
PredInt = data.frame(samp50, predict(sampMod, newdata=data.frame(gdpPercap=samp50$gdpPercap), 
    interval = "prediction"))

#Plot model with confidence and prediction bands
ggplot(PredInt, aes(x=log(gdpPercap), y=lifeExp)) + 
    geom_point() + 
    geom_smooth(method="lm", aes(fill="confidence"), alpha = 0.5) + 
    geom_ribbon(aes(y=fit, ymin=lwr, ymax=upr, fill="prediction"), alpha=0.2)
```

<br>
<br>


## More Practice with Models and Confidence Intervals


### Exercise 13

```{r}
yeargap <- gapminder %>% filter(year == 1952 | year == 2007)
```

```{r}
yeargap$year <- as.factor(yeargap$year)
```

```{r}
yeargap %>% ggplot(aes(y = lifeExp, x = log(gdpPercap), color = year)) + geom_smooth(method = 'lm')
```

```{r}
yearmod <- lm(lifeExp ~ log(gdpPercap) + year, yeargap)
summary(yearmod)
```

```{r}
#95% CI for 2007

7.8 + 2*0.3799
7.8 - 2*0.3799


-10.05+ 9.84 + 2*(2.95+0.9950)
-10.05+ 9.84 - 2*(2.95+0.9950)

#95% CI in 1952
-10.05 + 2*(2.95)
-10.05 - 2*(2.95)

#95% CI difference 2007-1952
 9.84 + 2*(0.9950)
 9.84 - 2*(0.9950)

```
Yes. When GDP is adjusted the year2007 has life expectancy increase of more than 0 years as compared to year1952
### Exercise 14
```{r}
AAgap <- gap2007 %>% filter(continent == 'Americas' | continent == 'Asia')
```

```{r}
AAgap %>% ggplot(aes(y = lifeExp, x = log(gdpPercap), color = continent)) + geom_smooth(method = 'lm')

```
```{r}
AAmod <- lm(lifeExp ~ log(gdpPercap) + continent, AAgap)
summary(AAmod)
```
```{r}
-1.4740 + 2*1.0938
-1.4740 - 2*1.0938
```

No. We don't have enough evidence to conclude that when controlling for GDP, life expectancies are significantly higher in the Americas than in Asia because 0 lies in the CI.




### Animation

```{r}
setwd('D:/Math2IndustryBootCamp2017')
for(i in seq(1952,2007,5)){
  #name = paste(as.character(i),'plot.png',sep='')
  #png(name)
  i
  gapminder %>% filter(year==i) %>% ggplot(aes(x = log(gdpPercap), y = lifeExp, color = continent, size = pop)) + geom_point()
  #dev.off()
  }

```



